Can Symmetric Cryptography Be Liberated from the Von Neumann Style?

### 2024-10-14 paper.tex
Typo

One solution\footnote{Needless to say, in reality, we first went
looking find 4 simplified implementations, inspired
by~\citet{adominicai:fixslicing-gift}. Only later did we use Coq to
understand how these solutions \emph{could} have been derived from
first principles.}, which consists in using Coq as a term rewriting
engine, is to postulate the existence of such an explicit form, let us
call it
%
\coqe!mixcolumns_mod_explicit!.
%
We then claim that this definition ought to be equivalent to its
specification, \coqe!mixcolumns_mod! for $n \in \{0, 1, 2, 3\}$.
Making sure that the operations of Boolean algebra, Naperian and
foldable functors are opaque, $\beta$ reduction yields a term in our
target language. However, it suffers from significant redundancy: by
working over the term algebra, we have not quotiented out the
equational theory. To do so, we orient the equational theory in a
rewrite database and apply the \coqe!autosubst! tactics to obtain
simplified forms. Similarly, we retrieve convenient \coqe!let! forms
thanks to the \coqe!set (ident := term)! tactics. We are then left
with reading off the definition of \coqe!mixcolumns_mod_explicit! from
the proof goal.

>>>>

One solution\footnote{Needless to say, in reality, we first went
looking for 4 direct, simplified implementations, inspired
by~\citet{adominicai:fixslicing-gift}. Then we showed that these
implementations were equivalent to their respective
specification. Only later did we use Coq to understand how these
solutions \emph{could} have been derived from first principles.},
which consists in using Coq as a term rewriting engine, is to
postulate the existence of such an explicit form, let us call it
%
\coqe!mixcolumns_mod_explicit!.
%
We then claim that this definition ought to be equivalent to its
specification, \coqe!mixcolumns_mod! for $n \in \{0, 1, 2, 3\}$.
Making sure that the operations of Boolean algebra, Naperian and
foldable functors are opaque, $\beta$ reduction yields a term in our
target language. However, it suffers from significant redundancy: by
working over the term algebra, we have not quotiented out the
equational theory. To do so, we orient the equational theory in a
rewrite database and apply the \coqe!autosubst! tactics to obtain
simplified forms. Similarly, we retrieve convenient \coqe!let! forms
thanks to the \coqe!set (ident := term)! tactics. We are then left
with reading off the definition of \coqe!mixcolumns_mod_explicit! from
the proof goal.
